---
title: "Final Exam"
subtitle: "DS-1000: Fall 2024"
author: "[YOUR NAME]"
institute: "Vanderbilt University"
date: "Due Date: 2024-12-10"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Overview

This is your final exam for DS-1000. It is an open-ended data science task, asking you to investigate an explanation of Donald Trump's support in the 2024 U.S. presidential election. It is cumulative in the sense that you are expected to apply concepts and skills learned over the course of the semester.

## Grading

The final exam is due by 11:59PM on Tuesday, December 10th. Two points will be deducted for each day late it is received. Submissions received after midnight on Thursday, December 12th will not be graded. 

Please upload **two** versions of this final. The first is a PDF of the **knitted** output, just like your problem sets which is used by the graders. The second is this .Rmd file in its raw form which is used by the professor to **apply a machine learning algorithm to check for violations of the honor code (see below)**. An additional 5 points will be deducted for failing to submit both files in the requested formats. 

## Resources

You are permitted to rely on any course resources from the Fall 2024 semester. These include all lecture slides, recordings, problem sets, answer keys, homeworks, ChatGPT (**copy the link in the space below**) and lecture notes, as well as any and all posts to Campuswire.

Campuswire access will be restricted during the week of the final exam You are only permitted to post clarifying questions about the exam, and these should only be made visible to the instructor and TAs. The graders, TAs, and the Professor will remove questions that ask for help on the contents of the exam.

## Honor

Unlike the problem sets, you are **prohibited** from working on this final together. You must digitally sign your name below, confirming that you did not collaborate on this exam with any of your classmates, share work, or otherwise discuss its contents. 

## Independent Work Statement
Please sign your name in the space provided by typing out your full name in place of the underline:

"I, __________________________________, am aware of the serious nature of plagiarism and affirm that I did not collaborate with other students while completing this final exam. I understand that violations of this agreement will result in a zero on the final exam, a failing grade for the semester, and a hearing with the Undergraduate Honor Council."

## ChatGPT Link [Optional]

*Copy the link to ChatGPT you used here: _______________________.

## Question 0
Require `tidyverse` and load the [`county_trump_2024.rds`](https://github.com/jbisbee1/DS1000_F2024/blob/main/data/county_trump_2024.rds?raw=true) data to an object called `countyTrump`. 


```{r}
# INSERT CODE HERE
```

## Codebook

The codebook for this dataset is produced below. Refer to this when interpreting regression coefficients!

| Name                  | Description                                                                                                                                                                                             |
|-----------------------|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------:|
| perc.trump.2024       | Proportion of all votes that were cast for Trump in 2024                                                                                                                                                |
| perc.trump.2020       | Proportion of all votes that were cast for Trump in 2020                                                                                                                                                |
| population            | County population in 2023                                                                                                                                                                               |
| perc.non.hisp.white   | Percent of the county that is non-Hispanic white in 2023                                                                                                                                                |
| perc.non.hisp.black   | Percent of the county that is non-Hispanic black in 2023                                                                                                                                                |
| perc.non.hisp.asian   | Percent of the county that is non-Hispanic asian in 2023                                                                                                                                                |
| perc.hispanic         | Percent of the county that is Hispanic in 2023                                                                                                                                                          |
| perc.male             | Percent of the county that is male in 2023                                                                                                                                                              |
| perc.65up             | Percent of the county that is 65 years or older in 2023                                                                                                                                                 |
| unemp.rate            | County unemployment rate (unemployed / in the labor force) in 2024                                                                                                                                      |
| lfpr                  | County labor force participation rate in 2024                                                                                                                                                           |
| weekly.wages          | Average weekly wages in the county in 2023                                                                                                                                                              |
| perc.rural            | The percent of the county that is classified as rural in 2020                                                                                                                                           |
| population.density    | The number of people per square mile in 2020                                                                                                                                                            |
| perc.manuf            | Percent of the county that is employed in manufacturing in 2023                                                                                                                                         |
| covid.death.rate      | Number of Covid-19 related deaths per 1,000 people in each county as of the fall of 2020                                                                                                                |



## Task [25 points]

*Please investigate these data and write up a report summarizing your findings. The only guidance is that we are interested in understanding the sources of Donald Trump's support in the 2024 election. You are free to choose whichever variables you think are most interesting.*

*The best reports will include the following information:*

*1. An introductory written response describing your research question, theory, and hypothesis, along with a verbal description of the data. [5 points]*

*2. Visualizations of the key variables of interest, and discussions any transformations you choose to apply, as well as visualizations of the key relationships of interest and a written summary of how to interpret these patterns. [5 points]*

*3. Estimation of a simple linear model of just $Y$ and $X$, along with a summary of the substantive interpretation of the relationships of interest, as well as evaluation of model fit. [5 points]*

*4. Estimation of an alternative linear model that controls for other variables, along with a discussion of why you chose the controls, and -- again -- summaries of the substantive interpretation and evaluation of model fit. [5 points]*

*5. A concluding written response summarizing your findings and highlighting shortcomings with your analysis that might be fixed with other data or other measures. [5 points]*

*6. EXTRA CREDIT: Identify the most important variable to add to the regression in the codebook, and discuss why it is so important. [5 points]*

---

