---
title: "Problem Set 2"
subtitle: "Mini Data Science"
author: "[YOUR NAME]"
institute: "Vanderbilt University"
date: "Due Date: 2024-09-06"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Getting Set Up

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps2.Rmd` to your `code` folder.

Copy and paste the contents of this `.Rmd` file into your `[LAST NAME]_ps2.Rmd` file. Then change the `author: [Your Name]` to your name.

We will be using the `sc_debt.Rds` file from the course [github page](https://github.com/jbisbee1/DS1000_F2024/blob/main/data/sc_debt.Rds).

All of the following questions should be answered in this `.Rmd` file. There are code chunks with incomplete code that need to be filled in. 

This problem set is worth 5 total points, plus 1 extra credit point. The point values for each question are indicated in brackets below. To receive full credit, you must have the correct code. In addition, some questions ask you to provide a written response in addition to the code.

You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments. To submit, compiled the completed problem set and upload the PDF file to Brightspace on Friday by midnight. Instructions for how to compile the output as a PDF can be found in [Problem Set 0](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/ds1000_pset_0.pdf) and in this [gif tutorial](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/save_as_pdf.gif). 

*Note that the TAs and professors will not respond to Campuswire posts after 5PM on Friday, so don't wait until the last minute to get started!*

**Good luck!**

*Copy the link to ChatGPT you used here: _________________



## Question 0 [0 points]
*Require `tidyverse` and load the `sc_debt.Rds` data by assigning it to an object named `df`.*
```{r}
require() # Load tidyverse
df <- read_rds() # Load the dataset
```


## Question 1 [1 point]
*Research Question: Do students who graduate from smaller schools (i.e., schools with smaller student bodies) make more money in their future careers? Before looking at the data, write out what you think the answer is, and explain why you think so.*

>- Write answer here

## Question 2 [1 point]
*Based on this research question, what is the outcome / dependent / $Y$ variable and what is the explanatory / independent / $X$ variable? What are their average values in the data?*

```{r}
df %>%
  summarise(..., # Summarise the average X variable
            ...) # Summarise the average of the Y variable
```

>- Write answer here

## Question 3 [1 point]
*Create the scatterplot of the data that analyzes your hypothesis, along with a line of best fit. Then, describe the result. Is your answer to the research question supported?*
```{r}
df %>%
  ggplot(aes(x = ..., # Put the explanatory variable on the x-axis
             y = ...)) +  # Put the outcome variable on the y-axis
  geom_...() + # Create a scatterplot
  geom_...(...) + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       x = '',
       y = '')
```

>- Write answer here.

## Question 4 [1 point]
*Does this relationship change by whether the school is a research university? Using the filter() function, create two versions of the plot, one for research universities and the other for non-research universities. What do you find?*

```{r}
df %>%
  filter(...) %>% # Filter to non-research universities
  ggplot(aes(x = ..., # Put the explanatory variable on the x-axis
             y = ...)) +  # Put the outcome variable on the y-axis
  geom_...() + # Create a scatterplot
  geom_...(...) + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       subtitle = '', 
       x = '',
       y = '')

df %>%
  filter(...) %>% # Filter to research universities
  ggplot(aes(x = ..., # Put the explanatory variable on the x-axis
             y = ...)) +  # Put the outcome variable on the y-axis
  geom_...() + # Create a scatterplot
  geom_...(...) + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       subtitle = '', 
       x = '',
       y = '')
```

>- Write answer here

## Question 5 [1 point]
*Instead of creating two separate plots, color the points by whether the school is a research university. To do this, you first need to modify the research_u variable to be categorical (it is currently stored as numeric). To do this, use the mutate command with `ifelse()` to create a new variable called `research_u_cat` which is either "Research" if `research_u` is equal to 1, and "Non-Research" otherwise.*
```{r}
df <- df %>%
  mutate(research_u_cat = ifelse(..., # Logic
                                 ..., # Value if true
                                 ...))# Value if false

df %>%
  ggplot(aes(x = ..., # Put the explanatory variable on the x-axis
             y = ...,      # Put the outcome variable on the y-axis
             color = ...)) + # Color by whether it is a research university
  geom_...() + # Create a scatterplot
  geom_...(...) + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       subtitle = '', 
       x = '',
       y = '')
```


## Extra Credit [1 points]
*Write a short paragraph discussing your findings. What do you think is going on in these data?*

>- Write answer here