---
title: "Problem Set 1"
subtitle: "Intro to `R`"
author: "[YOUR NAME]"
institute: "Vanderbilt University"
date: "Due Date: 2024-08-30"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Getting Set Up

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps1.Rmd` to your `code` folder.

Copy and paste the contents of this `.Rmd` file into your `[LAST NAME]_ps1.Rmd` file. Then change the `author: [Your Name]` to your name.

We will be using the `sc_debt.Rds` file from the course [github page](https://github.com/jbisbee1/DS1000_F2024/blob/main/data/sc_debt.Rds).

All of the following questions should be answered in this `.Rmd` file. There are code chunks with incomplete code that need to be filled in. 

This problem set is worth 5 total points, plus 1 extra credit point. The point values for each question are indicated in brackets below. To receive full credit, you must have the correct code. In addition, some questions ask you to provide a written response in addition to the code.

You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments. To submit, compiled the completed problem set and upload the PDF file to Brightspace on Friday by midnight. Instructions for how to compile the output as a PDF can be found in [Problem Set 0](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/ds1000_pset_0.pdf) and in this [gif tutorial](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/save_as_pdf.gif). 

*Note that the TAs and professors will not respond to Campuswire posts after 5PM on Friday, so don't wait until the last minute to get started!*

**Good luck!**

*Copy the link to ChatGPT you used here: _________________



## Question 0 [0 points]
*Require `tidyverse` and load the `sc_debt.Rds` data by assigning it to an object named `df`.*
```{r}
require(tidyverse) # Load tidyverse
df <- read_rds("https://github.com/jbisbee1/DS1000_F2024/raw/main/data/sc_debt.Rds") # Load the dataset
```

*The codebook is reproduced here.*

| Name           | Definition                                                                                                                                                                                             |
|----------------|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| unitid         | Unit ID                                                                                                                                                                                                |
| instnm         | Institution Name                                                                                                                                                                                       |
| stabbr         | State Abbreviation                                                                                                                                                                                     |
| grad_debt_mdn  | Median Debt of Graduates                                                                                                                                                                               |
| control        | Control Public or Private                                                                                                                                                                              |
| region         | Census Region                                                                                                                                                                                          |
| preddeg        | Predominant Degree Offered: Associates or Bachelors                                                                                                                                                    |
| openadmp       | Open Admissions Policy: 1= Yes, 2=No,3=No 1st time students                                                                                                                                            |
| adm_rate       | Admissions Rate: proportion of applications accepted                                                                                                                                                   |
| ccbasic        | Type of institution-- see [here](https://data.ed.gov/dataset/9dc70e6b-8426-4d71-b9d5-70ce6094a3f4/resource/658b5b83-ac9f-4e41-913e-9ba9411d7967/download/collegescorecarddatadictionary_01192021.xlsx) |
| selective      | Institution admits fewer than 10 % of applicants, 1=Yes, 0=No                                                                                                                                          |
| research_u     | Institution is a research university 1=Yes, 0=No                                                                                                                                                      |
| sat_avg        | Average Sat Scores                                                                                                                                                                                     |
| md_earn_wne_p6 | Average Earnings of Recent Graduates                                                                                                                                                                   |
|ugds  | Number of undergraduates |
| costt4a | Average cost of attendance (tuition-grants) |


## Question 1 [1 point]
*Which school has the highest future earnings and what is its admissions rate?*
```{r}
df %>% 
  arrange(desc(md_earn_wne_p6)) %>% # Arrange by the future earnings
  select(instnm,md_earn_wne_p6,adm_rate) # Select the school name, the admission rate, and the state
```

> - The University of Health Science and Pharmacy in St. Louis has the highest future earnings and an admissions rate of almost 92%. [Rubric: 0 points if no attempt. 0.5 points if correct written answer but code produces error. 0.5 points if correct code but no / wrong answer.]

## Question 2 [1 point]
*How many schools are located in Massachusetts*
```{r}
df %>%
  filter(stabbr == 'MA')
```

> - There are 91 schools in Massachusetts. [Rubric: 0 points if no attempt. 0.5 points if correct written answer but code produces error. 0.5 points if correct code but no / wrong answer.] 

## Question 3 [1 point]
*Create a new variable called `adm_rate_pct` which is the admissions rate multiplied by 100 to convert from a 0-to-1 decminal to a 0-to-100 percentage point. Then calculate the overall average admissions rate in the data and write this answer.*

```{r}
df <- df %>%
  mutate(adm_rate_pct = adm_rate*100)

df %>%
  summarise(adm_rate_pct_avg = mean(adm_rate_pct,na.rm=T))
```

> - The overall average admissions rate across all schools is 67.9%. [Rubric: 0 points if no attempt. 0.5 points if correct written answer but code produces error. 0.5 points if correct code but no / wrong answer. 0.75 points if code is correct but students forget `na.rm=T`, yielding an `NA`.]

## Question 4 [1 point]
*Calculate the average SAT score and median earnings of recent graduates by region. Which region has the highest average SAT scores and which has the highest future earnings?*
```{r}
df %>%
  group_by(region) %>% # Calculate region-by-region with group_by()
  summarise(sat_avg = mean(sat_avg,na.rm=T), # Summarise the average SAT
            earn_avg = mean(md_earn_wne_p6,na.rm=T)) %>% # Summarise the average earnings
  arrange(desc(sat_avg))
```

> New England has both the highest average SAT scores and the highest future earnings. [Rubric: 0 points if no attempt. 0.5 points if correct written answer but code produces error. 0.5 points if correct code but no / wrong answer. 0.75 points if code is correct but students forget `na.rm=T`, yielding an `NA`.]

## Question 5 [1 points]
*Research Question: Do students who graduate from smaller schools (i.e., schools with smaller student bodies) make more money in their future careers? Make sure to give an answer and then* **explain why** *you think so*.

>Yes students from smaller schools will make more money. This is because smaller schools tend to have smaller classes which means that professors can work with students directly, helping them learn faster and better.



## Extra Credit [1 point]
*Plot the average SAT score (x-axis) against the median earnings of recent graduates (y-axis) by school, and add the line of best fit. What relationship do you observe? Why do you think this relationship exists?*
```{r}
df %>%
  ggplot(aes(x = sat_avg,y = md_earn_wne_p6)) +  # Build the plot
  geom_point() + # Add the points
  geom_smooth(method = 'lm',se = F) # Add a line of best fit
```

> - I observe a positive relationship between SAT scores and earnings. I theorize that this relationship reflects the fact that SAT scores capture student abilities that are rewarded on the labor market. However, SAT scores are also correlated with many other socio-economic factors which might also improve one's earnings (i.e. social network) which are unrelated to student ability. [Rubric: 0 points if no attempt. 0.5 points if correct written answer but code produces error. 0.5 points if correct code but no / wrong answer. Written answers can vary, but should be logical.]

