---
title: "Problem Set 6"
subtitle: "Multivariate Visualization"
author: "[YOUR NAME]"
institute: "Vanderbilt University"
date: "Due Date: 2024-10-04"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Getting Set Up

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps6.Rmd` to your `code` folder.

Copy and paste the contents of this `.Rmd` file into your `[LAST NAME]_ps6.Rmd` file. Then change the `author: [Your Name]` to your name.

We will be using two datasets this week. The first is the `game_summary.Rds` file from the course [github page](https://github.com/jbisbee1/DS1000_F2024/blob/main/data/game_summary.Rds) and should be saved to an object called `games`. 

All of the following questions should be answered in this `.Rmd` file. There are code chunks with incomplete code that need to be filled in. To submit, compile (i.e., `knit`) the completed problem set and upload the PDF file to Brightspace on Friday by midnight. Instructions for how to compile the output as a PDF can be found in [Problem Set 0](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/ds1000_pset_0.pdf) and in this [gif tutorial](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/save_as_pdf.gif). 

This problem set is worth 5 total points, plus 1 extra credit point. The point values for each question are indicated in brackets below. To receive full credit, you must have the correct code. In addition, some questions ask you to provide a written response in addition to the code.

You will be deducted 1 point for each day late the problem set is submitted, and 1 point for failing to submit in the correct format.

You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments.

*Note that the TAs and professors will not respond to Campuswire posts after 5PM on Friday, so don't wait until the last minute to get started!*

**Good luck!**

*Copy the link to ChatGPT you used here: _________________

## Question 0
Require `tidyverse` and load the [`game_summary.rds`](https://github.com/jbisbee1/DS1000_F2024/blob/main/Lectures/4_Uni_Multivariate/data/game_summary.Rds?raw=true') data to an object called `games`.
```{r}
# INSERT CODE HERE
```


## Question 1 [1 point]
How many points, on average, did the Boston Celtics score at home and away games in the 2017 season? Calculate this answer and also plot the multivariate relationship. Explain why your chosen visualization is justified. Draw two vertical lines for the average points at home and away. (Gold star for those who can also add the label for the average points above each vertical line.)
```{r}
(vertLines <- games %>%
filter(...) %>% # Filter to the 2017 season AND to the Boston Celtics
  group_by(...) %>% # Group by the location of the game
  summarise(...)) # Calculate the average points

games %>%
  filter(...) %>% # Filter to the 2017 season AND to the Boston Celtics
  ggplot(aes(...)) + # Create a multivariate plot comparing points scored between home and away games
  geom_...() + # Choose the appropriate geom_... for this plot (i.e., geom_histogram(), geom_density(), geom_bar(), etc.)
  labs(...) + 
  geom_...(...) + # Add vertical lines for the average points scored at home and away.
  annotate(...) # Gold star: label the vertical lines
```

> Write answer here

## Question 2 [1 point]
Now recreate the same plot but show all three seasons side by side (hint: use `facet_wrap()`). Imagine that you work for the Celtics organization and Brad Stevens (the GM), asks you if the team scores more points at home or away? Based on your analysis, what would you tell him?

```{r}
# By season
(vertLines <- games %>%
filter(...) %>% # Filter to the Boston Celtics
  group_by(...) %>% # Group by the location and the season
  summarise(...)) # Calculate the average points

games %>%
  filter() %>% # Filter to the Boston Celtics
  ggplot(aes(...)) + # Create a multivariate plot comparing points scored between home and away games
  geom_...() + # Choose the appropriate geom_... for this plot (i.e., geom_histogram(), geom_density(), geom_bar(), etc.)
  labs(...) + 
  facet_wrap(...) + # Create separate panels for each season
  geom_vline(...) +  # Add vertical lines for the average points scored at home and away.
  geom_text(...) # Golder star: label the vertical lines AND COLOR THE TEXT
```

> Write answer here

## Question 3 [1 point]
Brad Stevens thanks you for your answer, but is a well-trained statistician in his own right, and wants to know how confident you are in your claim. Bootstrap sample the data 1,000 times to provide him with a more sophisticated answer. How confident are you in your conclusion that the Celtics score more points at home games than away games? Across 1,000 bootstrap simulations, what is the average point difference between home and away games? Make sure to `set.seed(123)` to ensure you get the same answer every time you `knit` your code!

```{r}
set.seed(...) # Set the seed to 123!
forBS <- games %>% # To make things easier, create a new data object that is filtered to just the Celtics
    filter(...) # Filter to the Celtics (nameTeam)

bsRes <- NULL # Instantiate an empty object to store data from the loop
for(i in 1:1000) { # Loop 1,000 times
  bsRes <- forBS %>%
    sample_n(...) %>% # Sample the data with replacement using all possible rows
    group_by(...) %>% # Group by the location of the game
    summarise(...) %>% # Calculate the average points
    ungroup() %>% # Best practices!
    mutate(... = ) %>% # Save the bootstrap index
    bind_rows(...) # Append the result to the empty object from above
} 

# Calculate the confidence
bsRes %>%
  pivot_wider(names_from = ...elt(),
              values_from = ...) %>% # Spread the data to get one column for average points at home and another for average points away
  mutate(...) %>% # Calculate the difference between Home and Away points
  summarise(..., # Calculate the proportion of bootstrap simulations where the home points are greater than the away points
            ...) # Calculate the overall average difference
```

> Write answer here

## Question 4 [1 point]

Visualize the bootstrapped results in two ways. First, plot the home and away simulations as densities with vertical lines for their averages, following the approach used in Questions 1 and 2 above. Second, plot the difference between home and away as a single density, with a vertical line at zero for reference.

```{r}
# You know the drill!
bsRes %>%
  ggplot(aes(...)) + 
  geom_...() + 
  geom_vline(...) + 
  labs(...)

bsRes %>%
  pivot_wider(names_from = ...,
              values_from = ...) %>% # Spread the data to get one column for average points at home and another for average points away
  mutate(...) %>% # Calculate the difference between Home and Away points
  ggplot(aes(...)) + 
  geom_...() + 
  geom_vline(...) +
  labs(...)
```


## Question 5 [1 point]
Re-do the analysis from question 3 for three other statistics of interest to Brad: total rebounds (treb), turnovers (tov), and field goal percent (pctFG). Do you notice anything strange in these results? What might explain it?

```{r}
set.seed(...) # Set the seed
bsRes <- NULL # Instantiate an empty object to store data from the loop
for(i in 1:1000) { # Loop 1,000 times
  bsRes <- forBS %>%
    sample_n(...) %>% # Sample the data with replacement using all possible rows
    group_by(...) %>% # Group by the location of the game (locationGame)
    summarise(..., # Calculate the average total rebounds (treb)
              ..., # Calculate the average turnovers (tov)
              ...) %>% # Calculate the average field goal shooting percentage (pctFG)
    ungroup() %>% # Best practices!
    mutate(...) %>% # Save the bootstrap index
    bind_rows(...) # Append the result to the empty object from above
} 

# Calculate the confidence
bsRes %>%
  pivot_wider(names_from = ..., # Pivot wider to get each measure in its own column for home and away games
                values_from = c(...)) %>% # Use the values from the variables you created above
    mutate(diff_reb = ..., # Calculate the difference between home and away total rebounds
           diff_tov = ..., # Calculate the difference between home and away turnovers
           diff_pctFG = ...) %>% 
  summarise(confidence_reb = ..., # Calculate the proportion of times the difference for rebounds is greater at home
            confidence_tov = ..., # Calculate the proportion of times the difference for turnovers is greater at home
            confidence_pctFG = ..., # Calculate the proportion of times the difference for FG% is greater at home
            avg_reb = ..., # Calculate the overall average home court advantage in rebounds
            avg_tov = ..., # Calculate the overall average home court advantage in turnovers
            avg_pctFG = ...) # Calculate the overall average home court advantage in FG%
```

> Write answer here

## Extra Credit [1 point]
Now Brad is asking for a similar analysis of other teams. Calculate the difference between home and away points for every team in the league and prepare a summary table that includes both the average difference for each team, as well as your confidence about whether the difference is not zero. Based on these data, would you argue that there is an **overall** home court advantage in terms of points across the NBA writ large? Visualize these summary results by plotting the difference on the x-axis, the teams (reordered) on the y-axis, and the points colored by whether you are more than 90% confident in your answer. How should we interpret confidence levels less than 50%?

```{r}
# INSERT CODE HERE (very hard)
```

> Write answer here

