---
title: "Lecture 15 Notes"
output: html_document
date: "2024-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting set up

1. Load `fn_cleaned_final.rds` from course GitHub

2. Look at the data

3. What predicts winning a game of Fortnite (`won`)?

```{r}
require(tidyverse)
fn <- read_rds('https://github.com/jbisbee1/DS1000_F2024/raw/main/data/fn_cleaned_final.rds')

glimpse(fn)

fn <- fn %>%
  rename(win = won)

glimpse(fn)

```

# Look at `hits`

```{r}
# Univariate visualization
# X: hits
fn %>%
  ggplot(aes(x = hits)) + 
  geom_histogram()

# Y: win
fn %>%
  ggplot(aes(x = win)) + 
  geom_bar()

```

# Multivariate VIsualization

```{r}
fn %>%
  ggplot(aes(x = hits,color = factor(win))) + 
  geom_density()

fn %>%
  ggplot(aes(x = factor(win),
             y = hits)) + 
  geom_boxplot()

# Bad idea: scatterplot
fn %>%
  ggplot(aes(x = hits,
             y = win)) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

# Add a second predictor: `mental_state`

```{r}
fn %>%
  ggplot(aes(x = mental_state)) + 
  geom_bar()

# Multivariate Visualization:
fn %>%
  ggplot(aes(x = mental_state,
             fill = factor(win))) + 
  geom_bar(position = 'fill') + 
  labs(y = 'Proportion of games')

```

# Trivariate visualization

- Introducing `geom_tile()`

```{r}
fn %>%
  group_by(mental_state,hits) %>%
  summarise(prob_win = mean(win)) %>%
  ggplot(aes(x = mental_state,
             y = factor(hits),
             fill = prob_win)) + 
  geom_tile()
```

# Introducing: `ntile()`

```{r}
fn %>%
  mutate(hits_decile = ntile(hits,n = 10)) %>%
  group_by(hits_decile,mental_state) %>%
  summarise(prob_win = mean(win)) %>%
  ggplot(aes(x = mental_state,
             y = factor(hits_decile),
             fill = prob_win)) +
  geom_tile()
```

# Simplest function: conditional means

- Calculate the average win by mental state

```{r}
fn %>%
  group_by(mental_state) %>%
  summarise(prob_win = mean(win))

# Step 1: use mutate() to add a column of predictions
fnTmp <- fn %>%
  group_by(mental_state) %>%
  mutate(prob_win = mean(win)) %>%
  select(mental_state,win,prob_win) %>%
  mutate(error = win - prob_win,
         pred_win = ifelse(prob_win > .25,
                           1,
                           0))

fnTmp %>%
  group_by(win,pred_win) %>%
  summarise(nGames = n())

fnTmp %>%
  count(prob_win)


# Calculate model's accuracy
fnTmp %>%
  group_by(win) %>%
  mutate(total_games = n()) %>%
  group_by(win,pred_win,total_games) %>%
  summarise(nGames = n()) %>%
  mutate(proportion_correct = nGames / total_games)
```

# Improving the model

```{r}
fn %>%
  group_by(mental_state,hits) %>% # Calculate probability of winning
  mutate(prob_win = mean(win)) %>%
  mutate(pred_win = ifelse(prob_win > .4, # Convert probabilities back to 0 and 1 values
                           1,
                           0)) %>%
  group_by(win) %>% # Calculate total games that were actually won or lost
  mutate(total_games = n()) %>%
  group_by(win,pred_win,total_games) %>% # Calculate model accuracy
  summarise(nGames = n()) %>%
  mutate(prop_correct = nGames / total_games) %>%
  ungroup() %>%
  mutate(prop_overall = sum(nGames*(win == pred_win)) / sum(nGames))

fn %>%
  summarise(mean(win))
```

# Throw it into a loop

```{r,message=FALSE,warning=FALSE}
threshRes <- NULL
for(thresh in seq(0,1,by = 0.025)) {
  tmp <- fn %>%
  group_by(mental_state,hits) %>% # Calculate probability of winning
  mutate(prob_win = mean(win)) %>%
  mutate(pred_win = ifelse(prob_win > thresh, # Convert probabilities back to 0 and 1 values
                           1,
                           0)) %>%
  group_by(win) %>% # Calculate total games that were actually won or lost
  mutate(total_games = n()) %>%
  group_by(win,pred_win,total_games) %>% # Calculate model accuracy
  summarise(nGames = n()) %>%
  mutate(prop_correct = nGames / total_games)
  
  threshRes <- threshRes %>%
    bind_rows(tmp %>%
                mutate(threshold = thresh))
}

threshRes
```

# Visualize the threshold results

```{r}
threshRes %>%
  mutate(metric = ifelse(win == 0 & pred_win == 0,"Specificity",
                         ifelse(win == 1 & pred_win == 1,"Sensitivity",
                                NA))) %>%
  drop_na(metric) %>%
  ggplot(aes(x = threshold,
             y = prop_correct,
             color = metric)) + 
  geom_line() + 
  geom_vline(xintercept = 0.32)
```

# Visualize the ROC Curve (Receiver Operator Characteristic)

```{r}
threshRes %>%
  mutate(metric = ifelse(win == 0 & pred_win == 0,"Specificity",
                         ifelse(win == 1 & pred_win == 1,"Sensitivity",NA))) %>%
  drop_na(metric) %>%
  ungroup() %>%
  select(prop_correct,metric,threshold) %>%
  pivot_wider(names_from = 'metric',
              values_from = 'prop_correct') %>%
  ggplot(aes(x = 1-Specificity,
             y = Sensitivity)) + 
  geom_line() + 
  geom_abline(slope = 1,intercept = 0,linetype = 'dotted') + 
  xlim(c(0,1)) + 
  ylim(c(0,1))
```

# Calculate AUC

```{r}
require(tidymodels)

# Create data for the roc_auc() function
forAUC <- fn %>%
  group_by(mental_state,hits) %>%
  mutate(prob_win = mean(win)) %>%
  ungroup()

# Convert true outcome to factor
forAUC <- forAUC %>%
  mutate(win = factor(win,levels = c('1','0')))

roc_auc(data = forAUC,
        win,
        prob_win)
```

# Trying a bunch of models

1. lm()
2. glm()
3. random forest

```{r}
# Linear regression
glimpse(fn)
m <- lm(win ~ mental_state + hits + player + eliminations + damage_taken,data = fn)

summary(m)

forAUC <- fn %>%
  mutate(prob_win = predict(m)) %>%
  select(prob_win,win) %>%
  mutate(win = factor(win,levels = c('1','0')))

roc_auc(data = forAUC,
        win,
        prob_win)

# Logistic regression
m2 <- glm(formula = win ~ mental_state + hits + player + eliminations + damage_taken,
          data = fn,
          family = binomial(link = 'logit'))

summary(m2)

# Calculate AUC
forAUC <- fn %>%
  mutate(prob_win = predict(m2,type = 'response')) %>%
  select(prob_win,win) %>%
  mutate(win = factor(win,levels = c('1','0')))

roc_auc(data = forAUC,
        win,
        prob_win)

```

