---
title: "Lecture 2 Notes"
output: html_document
date: "2024-08-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is a new section

## This is a smaller subsection

### This is a subsubsection

# Taking notes in RMarkdown

- The professor is demonstrating how to take notes

- He is typing too fast and making mistakes

- Or not

- This is **bold text**

- This is *italicized text*

- This is `code font`

# Coding in R

```{r}
2+2
```

# Demonstrating the object assignment operator `<-`

```{r}
object1 <- 2+2
```

blah blah blah filler texdt blah blah blah

```{r}
object1*2
```

# More complicated objects
```{r}
some_text <- "This is a sentence."
some_text <- 'This is a sentence.'


a_vector_of_numbers <- c(15,20,-15,101,32)
mean(a_vector_of_numbers)
median(a_vector_of_numbers)

rm(a_vector_of_numbers)

# Changing variable to something else
a_vector_of_numbers <- c(15,20,-15,101,32)
a_vector_of_numbers <- c("This","is","a","vector","of","words.")
a_mixture_vector <- c(125,"This",-1)

mean <- "This is a bad idea."

mean(c(15,315,3153,109531))

aMixture <- c(125,"This",-1)
```

# Using Packages

- Step 1: Install packages from the internet using `install.packages('tidyverse')`
- Step 2: Activate package using `require(tidyverse)`

```{r}
require(tidyverse)
df <- read_rds("https://github.com/jbisbee1/DS1000_F2024/raw/main/data/sc_debt.Rds")

df
```

# Two new functions:

- `filter()`
- `select()`

```{r}
df %>% filter(instnm == "Vanderbilt University") %>% select(md_earn_wne_p6)
```

